---
title: "Cat~Cat"
author: "Dr J"
date: "Mar 7, 2021"
output: html_document
---




# 2 x 2  Chisq or Fisher

This is a template to serve as an example (cut and paste, change variable names)  
we use the ucdavis1 dataset from the tigerstats package.

```{r include=FALSE}
require(tidyverse)
require(tigerstats)
require(knitr)
data(ucdavis1)
davisdata <- na.omit(ucdavis1)
```


### Plot barplots

#### Plot barplots tigerstats way

```{r}
barchartGC(~class + Sex,data=ucdavis1)
barchartGC(~class + Sex,data=ucdavis1, type="percent")

```

#### Plot barplots ggplot way

```{r}
dd2 <- davisdata %>% group_by(class,Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=class,y=count,fill=Sex))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=class,y=prcnt*100,fill=Sex)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
  
### Numerical Summaries  

```{r}
table1 <- xtabs(~class + Sex, data=davisdata)
rowPerc(table1)
colPerc(table1)
```
## Inferential  

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

# 2 x 3 (other than 2x2 case)


```{r}
dd3 <- davisdata %>% group_by(Seat,Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd3,aes(x=Seat,y=count,fill=Sex))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd3,aes(x=Seat,y=prcnt*100,fill=Sex)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

### Numerical Summaries  

```{r}
table2 <- xtabs(~Seat + Sex, data=davisdata)
rowPerc(table2)
colPerc(table2)
```

## Inferential  

```{r}
chisq.test(table2)
chisqtestGC(table2)
```

## A function to run multiple fisher tests
### Upload the fisherXY.R function then source it

```{r}
fisherXY = function(x, y, confL = 0, printAll=FALSE){
  # Inputs: 
  #   x: factor variable, 2 levels
  #   y: factor variable, n levels
  # confL: confidence level, default = 0.95
  #        print only those tables, results if p-value < (1-confL)
  # printAll: overide the 'print only if p-value < (1-confL), and
  #           print all tables and fisher test results
  #
  # Constraint:
  #   x and y must be same length
  #
  # Outputs:
  #   All 2x2 Fisher exact tests
  
  if(confL<0 || confL> 1){
    print('error: confL must be between 0 and 1')
    return(999)
  }
  
  if(!is.factor(x)) x=as.factor(x)
  if(!is.factor(y)) y=as.factor(y)
  
  fisherTable = as.matrix(table(x, y))
  cat('\ncontingency table:\n')
  print(fisherTable)
  cat('\n***All 2X2 tests****\n\n')

  # all 2x2 tests
  levelsY = levels(y)
  levelsX = levels(x)
  
  m = length(levelsX)
  n = length(levelsY)
  
  confLevel = 0.95
  if (confL !=0 ) confLevel=confL
  
  for (i in 1:(m-1))
       for(j in ((i+1):m))
          for (k in 1:(n-1))
             for (l in ((k+1):n)) {
                
                 fisherIndexX = c(i,j)
                 fisherIndexY = c(k,l)
                 
                 fTempTable = as.matrix(fisherTable[fisherIndexX,
                                        fisherIndexY])
                 
                 fisherTest=fisher.test(fTempTable, conf.int = TRUE, 
                              conf.level = confLevel)
                 
                 if((fisherTest$p.value < (1-confLevel) ) || printAll){
                   
                       print(c(levelsX[i], levelsX[j]))
                       print(c(levelsY[k], levelsY[l]))
                       print(fTempTable)
                       print(fisherTest)
                       cat('\n')
                 }
                 
        }
}

```
```{r}
fisherXY(davisdata$Sex,davisdata$Seat)
```

```{r}
FisherTable = function(cntngcyTable, confL = 0, printAll=FALSE){
  # Inputs:
  # cntngcyTable: mxn contingency table
  #
  # To make sense of the results, the table must have rownames and
  # column names.
  #
  #  Here's an example of how to set up a table
  #
  # suppose the research question were,
  # "Does councelling support affect pass/fail rates in math classes?"
  #
  # rows are 'See a counselor', 'Counselor strong recommendation',
  #          'Counselor recommendation', 'No Counselor Support'
  #
  # and the columns are 'failed' and 'passed'
  #
  #
  # the data look like:
  #                failed passed
  # SeeCounselor     11     17
  # StrongRec        53     28
  # Rec              97    100
  # NoSup           172    369
  #
  # To construct a data table for fisherTable:
  # use the following R code:
  #  >matMat = matrix(c(11,17,53,28,97,100,172,369), nrow=4,byrow = T)
  #  >colnames(matMat)=c("failed","passed")
  #  >rownames(matMat)=c("SeeCounselor","StrongRec", "Rec","NoSup")
  #  >mathTab = as.table(matMat)
  #  > print(mathTab)
  #
  #
  # confL: confidence level, default = 0.95
  #        print only those tables, results if p-value < (1-confL)
  #
  # printAll: overide the 'print only if p-value < (1-confL), and
  #           print all tables and fisher test results
  # Outputs:
  #   All 2x2 Fisher exact tests
 
  if(confL<0 || confL> 1){
    print('error: confL must be between 0 and 1')
    return(999)
  }
 
  cat('\ncontingency table:\n')
  print(cntngcyTable)
  if (printAll) {
    cat('\n***All 2X2 tests****\n\n')
  }else{
    cat('\n***Print only tests with significant results***\n\n')
  }
 
  # run through all 2x2 tests
  m = nrow(cntngcyTable)
  n = ncol(cntngcyTable)
 
  confLevel = 0.95
  if (confL !=0 ) confLevel=confL
 
  for (i in 1:(m-1))
    for(j in ((i+1):m))
      for (k in 1:(n-1))
        for (l in ((k+1):n)) {
         
          fisherIndexX = c(i,j)
          fisherIndexY = c(k,l)
         
          fTempTable = as.matrix(cntngcyTable[fisherIndexX,
                                   fisherIndexY])
         
          fisherTest=fisher.test(fTempTable, conf.int = TRUE,
                                 conf.level = confLevel)
         
          if((fisherTest$p.value < (1-confLevel) ) || printAll){
           
            print(fTempTable)
            print(fisherTest)
            cat('\n')
          }
        }
}
```

```{r ,eval=FALSE, include=F}
FisherTable(results,printAll = T)
```