---
title: "Eating Out"
author: "Shaun Ikeuchi"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`"    
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## There were two years in which the General Social Survey asked about the frequency of eating out. The variable name you are looking for is "eatout".
### Find those years and for the first year the question was asked download that years data in a stata format - its name will be "GSSxxxx.dta" where the xxxx is replaced by the year.  For example if the year was 1983 its name would be GSS1983.dta. Your assignment is to download the GSS yearly dataset for the first year the question was asked and then  write up the answer to the question by editing the  code block in line 31 to eval=TRUE and finish the job. 

$H_0$ There is no difference among the various races with frequency of eating out  
$H_A$ There is a statistical difference between at least two races with the frequency of eating out  


```{r include=FALSE,echo=FALSE}
source(file="fisherXY.R")
require(tidyverse)
require(tigerstats)
require(knitr)
require(foreign)
```
```{r, eval=FALSE}
GSS <-read.csv(file="GSS1996.csv")
GSS$race = as.factor(GSS$race)
GSS$eatout = as.factor(GSS$eatout)
GSS$race = case_match(GSS$race, '1'~'white', '2'~'black', '3'~"other")
GSS$eatout = case_match(GSS$eatout, '1'~'yes', '2' ~ 'no')

```

## Methods

'Race' and 'eatout' are categorical. Therefore, we use the cat~cat R markdown file.

### Graphical Description
## Plot BarPlots
```{r}


barchartGC(~eatout + race,data=GSS)
barchartGC(~eatout + race,data=GSS, type="percent")

```
 
The percentage of 
### Numerical Description

```{r}
table = xtabs(~eatout + race, data=GSS)
print(table)
rowPerc(table)
colPerc(table)
```

## Inferential

```{r}

chisq.test(table)
chisqtestGC(table)
fisher.test(table)

```
 
Since the p-value of the fischer test is 0.0003864 which is less than 0.05, it is statistically significant.
Since the p-value of the chi-squared test is 0.0008 which is less than 0.05, it is statistically significant. 

```{r}
dd2 <- GSS %>% group_by(eatout,race) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=race,y=count,fill=eatout))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=race,y=prcnt*100,fill=eatout)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

The number of white races eating out is signifcantly higher than black races eating out.


## Conclusions

There is a statistical difference the black and white race with the frequency of eating out. According to the bar chart, the number of white races eating out is 600 people comapred to the black population with only around 90 people. While the dataset of the black races is smaller, a large margin can still be seen. Furthermore, the p-value of the fischer test and the chi-squared test is less than 0.05, indicating that the null hypothesis can be rejected.  